### 2021-02-25 outputs/paper/paper.Rmd
Updated Appendix with survey screenshots and funding details

---
title: "TBD"
subtitle: "TBD"
author: "Najma Osman, Rachit Srivastava, Morgaine Westin, and Bongju Yoo"
thanks: "Code and data are available at: https://github.com/bonjwow/covid-shut-down"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2:
    fig_caption: yes
header-includes:
    #- \usepackage[labelformat=empty]{caption}
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bookdown)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(janitor)
library(dplyr)
library(palettetown)
library(RColorBrewer)
library(ggthemes)
library(scales)
library(broom)
library(purrr)
#devtools::install_github("johannesbjork/LaCroixColoR")
library(LaCroixColoR)
all_simulated_dataset <- 
  readr::read_csv("../../inputs/data/simulated_data.csv")
```

```{r, echo = FALSE, include=FALSE, message = FALSE}
#### Global Functions ####
### Format percent
percent <- function(x, decimal = 2, format = "f") {   
  paste0(formatC(x, format = format, digits = decimal), "%")
}
### Format table for Chi-squared test
tableChisq <- function(capt, stat, df, p) {  
  chi_stat <- c(stat)
  chi_param <- c(df)
  chi_p <- c(p) %>%
    format(scientific=FALSE)

  if (chi_p < 0.05 & chi_p >= 0.01) {
    astr <- "*"
  } else if (chi_p < 0.01 & chi_p >= 0.001) {
    astr <- "**"
  } else if (chi_p < 0.001) {
    astr <- "***"
  } else {
    astr <- ""
  }

  chi_df <- data.frame(chi_stat, chi_param, paste(sprintf("%1.3f", as.numeric(chi_p)), astr))

  kable(chi_df,
        digits = 3,
        caption = capt,
        booktabs = T,
        col.names = c("X-squared",
                      "df",
                      "p-value")) %>%
    footnote(general = "* p < 0.05, ** p < 0.01, *** p < 0.001",
             general_title = "") %>%
    column_spec(1, bold = T, width = "8em") %>%
    column_spec(2, width = "6em") %>%
    column_spec(3, width = "6em") %>%
    column_spec(4, width = "6em")
}
```
# Introduction

# Data

## Description of Study

We implemented a randomized control trial to examine the effects of shutting down indoor dining services on restaurant performance. Survey methodology was used to measure these effects against a valid comparison group consisting of restaurants operating at regular capacity. In this report, we will refer to the systematic closure of indoor dining services and simultaneously switch to take-out or delivery only services as the intervention. The control group will refer to restaurants whose services remained fully operational for the duration of the study and who were invited to complete a survey regarding their restaurant’s performance after the intervention period ended.

A sample size of 400 was selected for the experiment based on funding availability (see Appendix for funding details), with 200 restaurants randomly assigned to receive the intervention and 200 restaurants randomly assigned to be in the control group. While these sample sizes indicate the number of restaurants we reached out, the number of respondents who consented to participate in the study differed slightly. A total of 150 restaurants received the intervention and responded to the survey, while 180 restaurants assigned to the control condition completed the survey.

Restaurants randomly assigned to receive the intervention were asked to close their indoor dining services so their food was only available to customers for take-out or delivery. There were no restrictions as to how restaurants carried out their take-out and delivery orders, meaning they could use any type and number of food courier services, including their own in-house delivery services. Take-out included orders from individuals who called ahead or placed an order for pick-up through a food courier, as well as individuals who entered the store and made an order for take-out on the spot at the entrance. The only restriction was that seating areas in the restaurants were closed and customers had to leave with their food.

The duration of the intervention was three months, beginning on October 1st and ending on December 31st, 2020. While we understand that three months might be considered a long time to ask restaurants to partially shut down some of their services, we felt this length was necessary to ensure the intervention had time to stabilize and take effect. Since restaurant performance may fluctuate on a month-to-month basis due to a variety of external factors, we felt that measuring performance over a longer period of time would be a better indicator of the effects of the intervention and reduce the impacts of these fluctuations. The three month time period also gave restaurant owners and employees sufficient time to adjust and adapt to the new business model and make any changes where necessary. Finally, the duration of the intervention lined up with the third quarter of the fiscal year so revenue performance could be compared to the previous quarter. This was done purposefully under the assumption that many restaurants would already be calculating their quarterly earnings, and this would simplify their process of comparing performance before and after the intervention.

After 200 restaurants were randomly assigned to receive the intervention, the owners of each restaurant were contacted via telephone by members of the team at Petit Poll and invited to participate in the experiment. Restaurant owners were informed of the purpose of the intervention, the details and expected length of the intervention, and the compensation for participating. Restaurant owners were encouraged to ask questions and to ask for clarification when necessary to ensure transparency regarding the intervention process. On our end, this included making it clear that the restaurant may end up operating at a loss for the duration of the intervention, and that some employees may have to be put on a temporary leave of absence. In addition to this, restaurant owners were informed that they were under no obligation to participate in the experiment if they did not wish to do so, and that if they consented to participate, they were allowed to opt-out and revert to their standard business practice at any point during the intervention. Finally, while only the restaurant owner’s consent was necessary to participate in the experiment, they were strongly encouraged to discuss the intervention and its potential implications with their employees as well before deciding to participate. Restaurants were contacted 6 weeks before the starting date of the intervention to ensure an adequate amount of time to prepare for the transition. Informed consent forms signed by each restaurant owner were collected two weeks before the start date of the intervention. A total of 150 restaurants consented to participate in the intervention.

During the intervention period, members of the team at Petit Poll performed monthly check up calls with each restaurant in the intervention group. This was done to get a sense of how the intervention was going and allow restaurant owners to voice any concerns. This also allowed us to confirm if restaurants were still following the requirements of the intervention. Restaurant owners were once again reminded during these calls that they had the option to opt-out of the intervention at any time if they felt their business was at serious financial risk and did not wish to continue.

Restaurants in the intervention condition received $1200.00 CAD for participating. This amount was determined based on funding allocations for this study. We acknowledged that participating in the intervention may result in some financial losses and put some businesses at risk and wanted to provide some form of compensation to help alleviate those concerns as well as incentivize restaurants to participate. However, we also understand that this amount of money carries different weight for different types of restaurants. We provided resources and encouraged restaurants to look into external relief funding to support their businesses throughout and after the intervention period.

200 restaurants selected to be in the control condition were called or contacted via email after the end of the intervention and invited to complete a brief survey regarding various aspects of their restaurant’s performance, employees, and take-out and delivery practices. Restaurant owners were told the purpose of the study and invited to ask any questions they had. Like the intervention condition, restaurant owners were informed that their participation was voluntary, and their responses to the survey would be completely anonymized. A total of 180 restaurants assigned to the control condition completed the surveys within two weeks of the end of the intervention.

## Survey Details

A short survey hosted on Google Forms was created to assess the effects of the intervention. A link to the survey and screenshots of the questions can be found in the Appendix. The survey contained 16 questions, the majority of which were in categorical multiple choice format. Questions 1-4 in the survey were designed to gather basic demographic questions about the restaurants to characterize the sample, such as the restaurant’s location, scale, segmentation (e.g. fine dining, casual diving, or quick service), and the number of years the restaurant has been in operation. The rest of the questions were related to the intervention, and intended to measure a range of potential effects of the intervention on the restaurant. These included questions about the number of employees working at the restaurant before and after the intervention period, questions about whether or not any changes were made to menu prices, changes to hours or days of operation, and questions about the restaurants’ performance in terms of revenue. To avoid situations where respondents may be uncomfortable or not willing to share their exact revenue, questions were asked in relative terms to the last quarter. For example, question 7 asked restaurants to indicate whether their quarterly revenue was higher, lower, or around the same compared to the last quarter. Similarly, question 8 asked respondents’ to report how much their revenue declined compared to the last quarter in terms of percentages. These questions were both adapted from a survey conducted by Restaurants Canada on impacts of COVID-19 on businesses [@restcan].

Restaurants in the intervention and control groups received the exact same questions to ensure a fair comparison between groups. Two identical copies of the survey with separate links were created on Google Forms to differentiate between restaurants in the intervention group who completed the survey and those in the control group. To respect respondents’ privacy, restaurant owners within each group all received the same URL. Respondents were required to sign in with a google account to complete the survey to ensure only one response per restaurant, however their addresses were not saved or connected to their responses in any way to ensure complete anonymity. Only general demographic characteristics about the restaurant such as location and restaurant type were asked.

## Sampling Methodology

The restaurants included in the study were selected using stratified random sampling from a list of registered business licenses in Toronto issued by the Municipal Licensing and Standards Division. This dataset was obtained from the Toronto Open Data Portal [@todp2021] and contains information regarding the different types of licenses and permits issued to businesses in Toronto, their contact information, and location. Since all restaurants in Toronto need a license to operate, this dataset provides excellent coverage of all restaurants currently in the city. Additionally, since the dataset refreshes on a daily basis, all information including when businesses cancel their licenses is up to date. The data were filtered to keep only businesses categorized as eating establishments which had not cancelled their license at the time the dataset was accessed. Restaurants which had been identified as take-out only or had no seating accommodations in the dataset were also filtered out during the sampling process. This was done to help ensure any effects found after the experiment could be attributed to our intervention which required restaurants to operate on a take-out only basis. The inclusion of restaurants which have already been tailored as take-out only businesses before the start of the experiment might influence the results in unknown ways. This is due to the fact that these restaurants most likely already had strategies in place to optimize their business performance while only providing take-out.

Since we are interested in the effects of switching to a take-out only business model for all restaurants in the city of Toronto, we needed to make sure our sample was representative of restaurants all across the city and not just those clustered in certain areas. Similarly, we needed to ensure that there were no major regional differences between the restaurants who received the intervention and the restaurants in the control group. For example, if most of the restaurants who were selected to receive the intervention ended up being clustered in significantly different areas than those in the control group, a number of extraneous factors related to location could influence the results irrespective of the effects of the intervention. In order to mitigate these effects, the number of restaurants coming from each area in Toronto should be roughly equal between the intervention and control groups. Therefore, to ensure restaurants in each area of Toronto are accurately represented in the experiment and that the distribution of restaurants in the intervention and control groups, a stratified sampling method was used.

The first three characters of the postal codes provided for each restaurant in the dataset were used to stratify each restaurant into boroughs assigned by Canada Post [@canpost]. Figure \@ref(fig:regions) displays the percentage distribution of restaurants located in each borough. As expected, a large proportion of restaurants were located in Downtown Toronto, followed by North York and Scarborough. After assigning each restaurant to an individual strata, a stratified random sampling method was implemented in R [@citeR] to randomly select a number of restaurants from each borough in proportion to the number of restaurants clustered in each area. For example, since our total sample size is 400 and approximately 28% of restaurants in the total list of restaurants we are sampling from are located in Downtown Toronto, we randomly selected 28% of 400 (112 restaurants) from the Downtown Toronto area to be in our sample. After restaurants were selected from each borough, R was used to randomly assign each restaurant to either the intervention group or the control group. The random assignment was done by borough, so half of the restaurants selected from each borough were assigned to the intervention group and the other half were assigned to the control. This sampling process ensured the restaurants we randomly selected were representative of all areas in Toronto, and that there were no major differences between restaurants assigned to the intervention group and those assigned to the control group in terms of their locations.

```{r regions, echo = FALSE, message = FALSE, fig.cap="Distribution of Restaurants in Toronto by Canada Post FSA", fig.width=8, fig.height=3.6}
#### Pie Chart ####
### Import cleaned business licences dataset
dfCleanBiz <- readr::read_csv("../../inputs/data/clean_business_licences.csv")
### Create data frame
dfBoroughs <- data.frame(
  group = dfCleanBiz$fsa,
  value = dfCleanBiz$n
)
### Calculate percentages
dfBoroughs <- dfBoroughs %>% 
  arrange(desc(group)) %>%
  mutate(prop = value / sum(dfBoroughs$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop ) %>%
  arrange(desc(prop))
dfBoroughs <- dfBoroughs[-c(10),]
# print(dfBoroughs)
### Create pie chart
ggplot(dfBoroughs, aes(x="", y = prop, fill = group)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(x=1.65, y = ypos, label = percent(prop)), color = "black", size=2.8) +
  scale_fill_manual(values=lacroix_palette("PeachPear", n = 9)) + 
  labs(fill = "Region", title = "Proportion of Restaurants in Toronto by Borough") +
  theme(plot.title = element_text(face = "bold")) +
  theme_void() +
  theme(plot.title = element_text(face = "bold"))
```

In order for us to make causal inferences about the effects of our intervention, steps had to be taken to ensure we had a valid comparison group to compare the effects of the intervention to. One characteristic of a valid comparison group is that it has roughly the same characteristics as the intervention group [@gertler2016impact]. The Toronto restaurant scene is extremely diverse in terms of the size of the restaurants, the type of food served, and the target populations. While we wanted to cover a wide range of restaurant types, it was difficult to control for all these factors given the number of restaurants in Toronto. We relied on the process of randomized sampling and random assignment to ensure, at the very least, there were no systematic differences between groups since every restaurant had an equal chance of being assigned to either condition. Similarly, using stratification to control for location helped to ensure the groups were similar based on the number of restaurants picked from each borough. However, while the sampling method was selected to reduce differences between the groups, there may be significant differences between restaurants who choose to participate in the intervention versus those who do not which need to be addressed when discussing the effects of the intervention.

Finally, to make a valid comparison between the intervention and control groups, restaurants in each group needed to be unaffected by one another [@gertler2016impact]. To illustrate, consider a situation where one restaurant was selected to shut-down their indoor dining services and the restaurant located next door  was part of the control group and stays open. If a customer wanted to dine indoors at the first restaurant, sees that the restaurant is closed for indoor dining, and instead decides to eat at the restaurant next door because it is open, our results might be affected by the fact that the closure of one restaurant directly impacted the sales of another restaurant in the control group. Again, we relied on the random sampling methodology and the sample size in relation to the total number of restaurants in each area to circumvent this issue. Given the large number of restaurants we were sampling from, it is statistically unlikely that restaurants in the intervention and control groups would be so close together and impact each other in such a way.

## Results

Questions pertaining to the characteristics of the restaurants, such as location, size, type, and age of restaurants (Figures 1-5) show us that:

* the largest concentration (nearly 60%) of restaurants in the Greater Toronto Area are centralized in downtown Toronto (28.26%), North York (15.78%), and Scarborough (13.74%);
* most restaurants are of medium or larger scale, with seating areas of at least 30; 
* a majority of the restaurants are relatively new (3 years or less); 
* fast-food/quick service restaurants are most common, followed by casual dining, and then fine dining;
* and UberEats is the most popular food courier service in the city.

Tables 1 and 2 speak to changes in numbers of employees for restaurants prior to the intervention period (Oct 1) and following the intervention period (Dec 31). This shows us that initially, both the control and treatment group had a very similar number of employees on average. Following the intervention period, however, the treatment group appears to have 10.26 fewer employees on average. Two-tailed Welch's t-tests comparing the two groups' mean employees shows that the difference prior to the intervention period is not a significant finding ( *t*(330) = -8.56, *p* > .05)., but the difference in employee numbers between the two groups post-intervention is significant( *t*(330) = 19.46, *p* < .05).

In terms of effects on revenue and/or overall sales for restaurants during the intervention period, Figure 5 shows us that more restaurants within the treatment group saw decreases in revenue compared to the prior quarter (July 1 - Sep 30), while the control group's revenue largely remained the same. In addition to this, Figure 6 shows that of those restaurants that saw a decrease in revenue, the treatment group saw the harshest declines in revenue during the intervention period. A Chi-Square test for whether the restaurant saw a decrease in revenue showed that restaurants in the treatment group were more likely to report a decrease in revenue *X*^2^ (2, *N* = 330) = 220.73, *p* < 2.2e-16. The percentage decrease in revenues also showed a significant relationship, where restaurants in the treatment group are most likely to report higher loss of revenue during the intervention period *X*^2^ (3, *N* = 330) = 178.33, *p* < 2.2e-16. Prior to the intervention period, it appears as though both groups had similar breakdowns of revenue coming from take-out and/or delivery (Figure 7).

Questions related to the adjustments of the restaurants during the intervention, such as a menu price, and hours and days of operation (Figures 8-12) show us that:

* most of the restaurants in the control group did not change the menu price, while most of the ones in the treatment group either lowered or raised the menu price;
* and most of the restaurants in the control group did not adjust their hours and days of operation, while most of the ones in the intervention group reduced their operating schedule.

Pearson’s Chi-squared (*X*^2^) test is used to see if there is a statistically significant difference between the intervention and the adjustments of the restaurants. The test results show that the p-value of each of those adjustments is less than 0.001, meaning that there is a high correlation between those variables. Consequently, it is expected that restaurants will most likely adjust the menu price and reduce their operating schedule if they operate on a delivery and take-out only basis.

## Figures

### Sample Characteristics

```{r Q1, echo = FALSE, message = FALSE, fig.cap = "Location of restaurants by borough"}
ggplot(data = all_simulated_dataset, mapping = aes (x = Q1, fill = type)) +
  geom_bar() + coord_flip() +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  labs(fill = "Condition", title = "Location of restaurants")
```

```{r Q2, echo = FALSE, message = FALSE, fig.cap = "Scale of restaurant in terms of space and seating capacity"}
### How would you best describe the scale of your restaurant? (in terms of space and seating capacity)
### Small (full seating capacity is under 30)
### Medium (full seating capacity between 30 and 60)
### Large (full seating capacity is 60+) "
ggplot(data = all_simulated_dataset, mapping = aes (x = Q2, fill = type)) +
    geom_bar() +
  scale_x_discrete(limit = c ("Small", "Medium", "Large")) +
  labs(x = "Seating Capacity",
       y = "# of Restaurants",
       fill = "Condition",
       title = "Scale of restaurant") +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))
```

```{r Q3, echo = FALSE, message = FALSE, fig.cap = "Types of Restaurants"}
### What best describes your restaurant segmentation?
### Fast-food/quick service
### Casual Dining
### Fine Dining
### Other
ggplot(data = all_simulated_dataset, mapping = aes (x = Q3, fill = type)) +
  geom_bar() +
  scale_x_discrete(limit = c ("Fine Dining", "Casual Dining", "Fast-food/quick service", "Other")) +
  labs(x = "Restaurant Type",
       y = "# of Restaurants",
       fill = "Condition",
       title = "Type of restaurant") +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))
```

```{r Q4, echo = FALSE, message = FALSE, fig.cap = "Number of years restaurants have been in operation"}
### How many years has your restaurant been in operation?
### 1-4 years
### 5-7 years
### 8-10 years
### 10 or more years
ggplot(data = all_simulated_dataset, mapping = aes (x = Q4, fill = type)) +
  geom_bar() + coord_flip() +
  scale_x_discrete(limit = c ("3 years or less", "5-7 years", "8-10 years", "10 or more years")) +
  labs(x = "Years of Operation",
       y = "# of Restaurants",
       fill = "Condition",
       title = str_wrap("Number of years in operation", 30)) +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))
```

```{r Q10, echo = FALSE, message = FALSE, fig.cap = "The most used food courier service"}
### Between Oct 1 and Dec 31, which food courier service did you receive the most orders from?
### UberEats
### SkiptheDishes 
### DoorDash
### Other
### In-house delivery service
### None
ggplot(data = all_simulated_dataset, mapping = aes(x = Q11, fill = type)) +
  geom_bar(position = "stack") + coord_flip() +
  scale_x_discrete(limit = c("UberEats", "SkiptheDishes", "DoorDash", "In-house delivery service", "Other", "None")) +
  labs(x = "Delivery Service",
       y = "# of Restaurants",
       title = "Food courier service popularity",
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) + 
  theme_fivethirtyeight()
```

### Effects of Intervention on Number of Employees

```{r Q5Q6mean, echo = FALSE, message = FALSE}
### Prior to (intervention start date), how many people did you employ?
### How many people do you currently employ?
t1 <-
all_simulated_dataset %>%
  group_by(type) %>%
  summarise(prior_employees = mean(Q5, na.rm = TRUE, .groups = NULL))
t2 <-
all_simulated_dataset %>%
  group_by(type) %>%
  summarise(curr_employees = mean(Q6, na.rm = TRUE), .groups = NULL)
total <- merge(t1, t2, by = c("type"))
# kable(total,
#       booktabs = T,
#       digits = 3,
#       col.names = c("Group Type",
#                   "# of Employees Prior to Intervention",
#                   "Current # of Employees")) %>%
#       footnote(general = "Change in # of employees before and after intervention period (Q5 + Q6)") %>%
#   kable_paper()
  # This table could be excluded because the means will be displayed anyway in the following T-test table?
```

```{r Q5Q6tTest, echo = FALSE, message = FALSE}
### T-Tests (Welch's)
#Q5
control_q5 <-
  all_simulated_dataset %>%
  filter(type == "Control") %>%
  select(Q5)
treatment_q5 <-
  all_simulated_dataset %>%
  filter(type == "Treated") %>%
  select(Q5)
t3 <- t.test(control_q5, treatment_q5, paired = FALSE)
#Q6
control_q6 <-
  all_simulated_dataset %>%
  filter(type == "Control") %>%
  select(Q6)
treatment_q6 <-
  all_simulated_dataset %>%
  filter(type == "Treated") %>%
  select(Q6)
t4 <- t.test(control_q6, treatment_q6, paired = FALSE)

### Q5 + Q6 compiled t-test table
t_tests <- map_df(list(t3, t4), tidy) %>%
  subset(select = -c(alternative, method, conf.low, conf.high))

# Round values
t_tests[] <- lapply(t_tests, round, 3)

# Output data table
t_tests %>%
  kable(
    booktabs = T,
    caption = "Welch's T-test to examine effects of intervention on the number of employees",
    digits = 3,
    col.names = c("Mean Difference",
                  "Mean (Pre-Intervention)",
                  "Mean (Post-Intervention)",
                  "t",
                  "p",
                  "df")

  ) %>%
  add_header_above(c(" ", "Control" = 1, "Treated" = 1, " ", " ", " ")) %>%
  kable_styling(full_width = F)
```

### Effects of Intervention on Revenue

```{r Q7, echo = FALSE, message = FALSE}
### Between Oct 1 and Dec 31 was your total sales revenue higher, lower, or about the same as it was during the previous quarter (last quarter dates)? *
### Higher
### Lower
### Around the Same
ggplot(data = all_simulated_dataset, mapping = aes(x = Q7, fill = type)) + 
  geom_bar() +
  labs(x = "Revenue Change",
       y = "# of Restaurants",
       title = "Revenue changes between Oct 1 and Dec 31",
       fill = "Condition") +
  theme_fivethirtyeight() + 
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))

### Chi-squared test
q7_chi <- table(all_simulated_dataset$Q7, all_simulated_dataset$type) %>%
  chisq.test()
tableChisq("Pearson's Chi-squared test to examine the possibility of an association between the intervention and the total sales", q7_chi$statistic, q7_chi$parameter, q7_chi$p.value)
```

```{r Q8, echo = FALSE, message = FALSE}
### What was the approximate percentage decrease in total sales revenue between Oct 1 and Dec 31 compared to the during the previous quarter (last quarter dates)? *
### Decline of 50% or more
### Decline of 25% to 50%
### Decline of 10% to 25%
### Decline of 10% or less
ggplot(data = all_simulated_dataset, mapping = aes (x = Q8, fill = type)) +
  geom_bar(position = "stack") + coord_flip() +
  scale_x_discrete(limit = c("Decline of 50% or more", "Decline of 25% to 50%", "Decline of 10% to 25%", "Decline of 10% or less")) + 
  labs(x = "Aprrox. % Decrease in Total Sales Revenue",
       y = "# of Restaurants",
       title = str_wrap("Percentage of total sales revenue decrease between Oct 1 and Dec 31", 40),
       fill = "Condition") +
  theme_fivethirtyeight() + 
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))

### Chi-squared test
q8_chi <- table(all_simulated_dataset$Q8, all_simulated_dataset$type) %>%
  chisq.test()
tableChisq("Pearson's Chi-squared test to examine the relation between the intervention and the total sales and a decline in revenue", q8_chi$statistic, q8_chi$parameter, q8_chi$p.value)
```

```{r Q9, echo = FALSE, message = FALSE}
### Before Oct 1st, approximately how much of your total monthly sales revenue came from takeout or delivery? *
### Did not offer takeout or delivery
### 25% or less
### 25- 50%
### 50-75%
### 75% or more

my_title = "Total revenue from takeout/delivery prior to Oct 1"
ggplot(data = all_simulated_dataset, mapping = aes (x = Q9, fill = type)) +
  geom_bar(position = "stack") +  coord_flip() + 
  scale_x_discrete(limit = c("Did not offer takeout or delivery", "25% or less", "25-50%", "50-75%", "75% or more")) + 
  labs(x = "Total Monthly Sales Revenue Fromt Takeout or Delivery",
       y = "# of Restaurants",
       title = str_wrap("Total revenue from takeout or delivery prior to Oct 1", 30),
       fill = "Condition") +
  theme_fivethirtyeight() + 
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))
```

## Effects of Intervention on Menu Prices

```{r Q11, echo = FALSE, message = FALSE}
### Between (start date of intervention) and (end date of intervention), did you make any adjustments to your menu prices?
### Yes, prices were raised
### Yes, prices were lowered
### No changes were made
ggplot(data = all_simulated_dataset, mapping = aes (x = Q10, fill = type)) +
  geom_bar() +  
  coord_flip() + 
  scale_x_discrete(limit = c("Yes, prices were raised", "Yes, prices were lowered", "No changes were made")) + 
  labs(x = "Adjustments to Menu Prices",
       y = "# of Restaurants",
       title = str_wrap("Menu adjustments between Oct 1 and Dec 31", 30),
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight()

### Chi-squared test
q11_chi <- table(all_simulated_dataset$Q11, all_simulated_dataset$type) %>%
  chisq.test()
tableChisq("Pearson's Chi-squared test to examine the association of the intervention with price adjustments", q11_chi$statistic, q11_chi$parameter, q11_chi$p.value)
```

## Effects of Intervention on Hours/Days of Operation

```{r Q12, echo = FALSE, message = FALSE}
### Between (start date of intervention) and (end date of intervention) did you reduce your ### restaurant’s hours of operation?
### Yes
### No
ggplot(data = all_simulated_dataset, mapping = aes(x = Q12, fill = type)) +
  geom_bar(position = position_dodge(1)) +
  labs(x = "Reduction in hours of operation",
       y = "# of restaurants",
       title = str_wrap("Reducton in operating hours between Oct 1 and Dec 31", 30),
       fill = "Condition") +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))

### Chi-squared test
q12_chi <- table(all_simulated_dataset$Q12, all_simulated_dataset$type) %>%
  chisq.test()
tableChisq("Pearson's Chi-squared Test for Q12", q12_chi$statistic, q12_chi$parameter, q12_chi$p.value)
```

```{r Q13, echo = FALSE, message = FALSE}
### If yes, approximately how many hours per week did you reduce by
### 15% or less
### 15-30%
### 30-50%
### 50-75%
### 75% or more
ggplot(data = all_simulated_dataset, mapping = aes(x = Q13, fill = type)) +
  geom_bar(position = "stack") +
  labs(x = "Hours Reduced per Week",
       y = "# of Restaurants",
       title = str_wrap("Percentage of hours reduced between Oct 1 and Dec 31", 30),
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight()
```

```{r Q14, echo = FALSE, message = FALSE}
### Between (start date of intervention) and (end date of intervention) did you reduce the 
### number of days your restaurant was open to the public?
### Yes
### No 
ggplot(data = all_simulated_dataset, mapping = aes(x = Q14, fill = type)) +
  geom_bar(position = position_dodge(1)) +
  labs(x = "Reduction in Days of Operation",
       y = "# of Restaurants",
       title = str_wrap("Reduction of operating days between Oct 1 and Dec 31", 30),
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight()

### Chi-squared test
q14_chi <- table(all_simulated_dataset$Q14, all_simulated_dataset$type) %>%
  chisq.test()
tableChisq("Pearson's Chi-squared Test for Q14", q14_chi$statistic, q14_chi$parameter, q14_chi$p.value)
```

```{r Q15, echo = FALSE, message = FALSE}
### If yes, how many days per week did you reduce by?
### 1-2 days
### 3-4 days
### 5-6 days
### 7 days
### Not Applicable
### bar graph/pie chart/line
ggplot(data = all_simulated_dataset, mapping = aes(x = Q15, fill = type)) +
  geom_bar(position = "stack") +
  labs(x = "Days Reduced",
       y = "# of Restaurants",
       title = "Reduction in days of operation per week",
       fill = "Condition") +
    scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight()
```

```{r Q16, echo = FALSE, message = FALSE}
### Have you applied for and received any external funding or grants for your restaurant since (intervention start date)?
### Yes
### No
ggplot(data = all_simulated_dataset, mapping = aes(x = Q16, fill = type)) +
  geom_bar(position = position_dodge(1)) +
  labs(x = "Application status",
       y = "# of restaurants",
       title = str_wrap("Funding Applications between Oct 1 and Dec 31", 30),
       fill = "Condition") +
  theme_fivethirtyeight() +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE))
```

\newpage

# Discussion

## Overview of Findings

## Limitations and Future Directions

\newpage

# Appendix {-}

## Survey

```{r survey1, echo = FALSE, message = FALSE, fig.cap = "Screenshot of Google Forms Survey Questions 1-8", out.width="90%"}
knitr::include_graphics("survey/01.png")
```

```{r survey2, echo = FALSE, message = FALSE, fig.cap="Screenshot of Google Forms Survey Questions 9-16", out.width="90%"}
knitr::include_graphics("survey/02.png")
```

\newpage

## Funding Acknowledgement

We would like to thank the Ontario Government for generously providing Petit Poll with $240,000 CAD in funding to carry out this study. The funding was used to provide compensation to restaurants for participating in the intervention. A sample size of 200 for the intervention condition was selected based on this amount, in which each restaurant who participated in the intervention received $1200. A total of 150 restaurants participated in the intervention and received payment, making the total amount of funding spent during the study $180,000 CAD.

\newpage

# References